---
title: "R Reference Guide"
author: "for Madison Brady"
date: "Fall 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Welcome to Madison's course page for ECO 602/634!
[Final Project: R Reference Guide](final_R_reference.html)

## R Reference Guide {.tabset .tabset-pills}

### Loading Data and Packages
1. **install.packages()** is used to install new functions for R

2. **require(name of function)** is used to bring the desired function/a new file into the current working directory, and must be used after installing a package in order to make it work

3. **read.csv(file name on computer)** is used to read a file in csv format from the working directory

4. **library()** can tell you the file path's name that you are working on, which can then be used with install.packages to tell R exactly where to save a package

5. **here()** is a downloaded function, which must be require()'d in order to use it in your working directory


**example:**

step 1: download a data file into the data folder 

step 2: read the data file and download it into an R session:

```{r}
library(readr)
ginkgo_data_2021 <- read_csv("C:/Users/madis/Desktop/environmental_data - Copy/data/ginkgo_data_2021.csv")
ginkgo <- data.frame(ginkgo_data_2021)
```




### Distribution Functions
**For a discrete parametric distribution, the distribution functions allow you to ask the following questions:**

1. What is the probability that I observe a value of exactly x?: The probability mass function, e.g. dbinom()

2. What is the probability that I observe a value of x or less?: The cumulative mass function, e.g. pbinom()

3. What is the probability that I observe a value of x or more?: The cumulative mass function, e.g. pbinom()

4. What is the median or 50th percentile?: The quantile function, e.g. qbinom()

5. What is the 90th percentile?: The quantile function, e.g. qbinom()

**For a continuous parametric distribution, the distribution functions allow you to ask the following questions:**

1. Is a value of 1.2 or 2.4 more likely? - The probability density function, e.g. dnorm()

2. What is the probability that I observe a value between 1.2 and 2.4?: The cumulative density function, e.g. pnorm()

3. What is the probability that I observe a value of 1.3 or more?: The cumulative density function, e.g. pnorm()

4. What is the probability that I observe a value of 2.4 or less?: The cumulative density function, e.g. pnorm()

5. What is the 20th percentile of fish lengths?: The quantile function, e.g. qnorm()

**Reminder** sum of all probabilities must equal **1**. The answer may be equivalent to 1-pnorm(), etc.



### Data Structures
1. **c()** combines values into a vector or list

2. **length()** gets or sets the length of vectors/lists and factors

3. **matrix()** creates a matrix from a given set of values

4. **data.frame()** create a data frame using existing vectors/lists. Data frames are used as the fundamental data structure by most of R's modeling software


**examples:**

```{r}
nrow(ginkgo)
ncol(ginkgo)
dim(ginkgo)
```

Conclusion: the ginkgo data file has 756 rows, 6 columns, and the dimensions 756 cells by 6 cells




### Subsetting
1. **subset()** returns subsets of vectors, matrices or data frames which meet specified conditions (uses Boolean/logical phrases)

2. or use **$** to select a specific column from a data set or data frame:

3. or use **[ ]** to select a subset of data based on row/column:

all rows/columns - df[,]


first row and all columns - df[1,]


first two rows and all columns - df[1:2,]


first and third row and all columns - df[ c(1,3), ]


first row and second and third column - df[1, 2:3]


first and second row and second and third column - df[1:2, 2:3]


**examples:**

```{r}
petiole <- ginkgo$petiole_length
petiole

library(readr)
penguins <- read_csv("C:/Users/madis/Desktop/environmental_data - Copy/data/penguins.csv")

Adelie <- subset(penguins, species == "Adelie")
Adelie

first.row.all.columns <- ginkgo[1,]
first.row.all.columns

row.2.column.3 <- ginkgo[2,3]
row.2.column.3

all.rows.column.3 <- ginkgo[,3]
all.rows.column.3
```



### Numerical Data Exploration

1. **summary()** - prints a results summary 

2. **mean()** - prints out the mean of a vector

3. **sd()** - prints out the standard deviation of a vector

**examples**
```{r}
summary(ginkgo)

mean(na.omit(petiole))

sd(na.omit(petiole))
```


### Graphical Data Exploration
1. **par(mfrow = c(1,2))** is a function that tells the plot window in R how to arrange a series of figures. In this instance, **c(1,2)** asks for a plot window with 2 figures arranged in **1 row** with **2 columns**


2. **plot()** required arguments:

col = color of points

pch = type of point indicated by a number (Google to see a key)

cex = magnification extent in relation to default sizing

main = title (must be in quotes)

sub = subtitle (must be in quotes)

xlab = x axis label (must be in quotes)

ylab = y axis label (must be in quotes)

xlim = max value for x axis

ylim = max value for y axis

type = "p" for point (default), "l" for line, "b" for both

to add a line, use abline() or use plot() again with the added specification that add = TRUE


3. **hist()** required arguments are mostly same as plot() function, with the addition of:

breaks = numeric, indicates distance between bars/placement on the x axis


4. **boxplot()** required arguments are mostly the same as plot() function, with the addition of:

boxplot = subset ~ subset, data set to specify how data is organized into different boxes


**examples:**

```{r}
plot(ginkgo$max_depth, ginkgo$max_width, main = "Ginkgo Leaf Max Width vs. Max Depth", xlab = "max depth (mm)", ylab = "max width (mm)", pch = 5, type = "p", col = "red")
```
```{r}
hist(penguins$flipper_length_mm, main = "Histogram of Flipper Lengths", xlab = "flipper length (mm)", ylab = "Frequency", col = "blue")
```
```{r}
boxplot(ginkgo$petiole_length, main = "Boxplot of Petiole Length", xlab = "ginkgo leaves", ylab = "petiole length (mm)", col = "yellow")
```

```{r}
boxplot(petiole_length ~ seeds_present, data = ginkgo, main = "Boxplot of Petiole Length by Seeds Present", xlab = "seeds present?", ylab = "petiole length (mm)", col = "yellow3")
```

```{r}
par(mfrow = c(2,2))
hist(ginkgo$max_width, main = "Max Width of Ginkgo Leaves", xlab = "width (mm)", col = "cyan1")
hist(ginkgo$max_depth, main = "Max Depth of Ginkgo Leaves", xlab = "depth (mm)", col = "cyan2")
hist(ginkgo$notch_depth, main = "Max Notch Depth of Ginkgo Leaves", xlab = "notch depth (mm)", col = "cyan3")
hist(ginkgo$petiole_length, main = "Petiole Length of Ginkgo Leaves", xlab = "petiole length (mm)", col = "cyan4")
```







